As-you-go" instead of "after-the-fact": A network approach to scholarly communication and evaluation

### 2018-03-30 docs/manuscript.Rmd
Intro done by Chris

  Scholarly research faces threats to its sustainability and has been said to face a reproducibility crisis [@Baker2016] amongst other pernicious problems such as access and exclusivity. The underlying cause might be the way we have collectively designed the reporting and rewarding of  research (implicitly or explicitly). The current scholarly communication system is primarily organized around researchers who publish static (digital) research papers in journals. Many of these journals have artificial page limits (in the digital age), which leads to artificial scarcity and subsequently increases the perceived prestige of such a journal due to high rejection rates (71% on average for APA journals in 2016; [perma.cc/Q7AT-RN5C](https://perma.cc/Q7AT-RN5C)). Furthermore, scholarly communication has become highly centralized, where over 50% of all papers are published by as little as five publishers (over 70% for social sciences) [@Lariviere2015]. Centralization has introduced knowledge discrimination, as publishers are able to influence who can access scholarly knowledge, what gets published, and allows for other single points of failure to arise with their own consequences (e.g., censorship; https://perma.cc/HDX8-DJ8F). In order to have a sustainable scholarly research system, we consider it necessary to implement changes that provide progress on multiple of these threats at once instead of addressing them individually.

  Systems design directly affects what the system and the people who use it can do; scholarly communication still retains an analog based design affecting the effectivity of the spread and production of knowledge dissemination [see also @10.1002/aris.1440370105]. Researchers and institutions are evaluated on where and how much papers they publish (as a form of prestige). For example, an oft-used measure of quality is the Journal Impact Factor (JIF) [@Garfield2006]. The JIF is also frequently used to evaluate the 'quality' of individual papers under the assumption that a high impact factor predicts the success of individual papers (this assumption has been debunked many times) [@Prathap2016; @Seglen1992; @Seglen1994]. Many other performance indicators in the current system (e.g., citation counts and h-indices) resort to generic bean counting. Inadequate evaluation measures leave universities, individual researchers, and funders (amongst others) in the dark with respect to the substantive questions they might have about the produced scholarly knowledge. Additionally, work that is not aptly captured by the authorship of papers is likely to receive less recognition (e.g., writing software code) due to reward systems counting publications instead of contributions (see also https://perma.cc/MUH7-VCA9). It is impossible that a paper-based approach to scholarly communication can escape the consequences of paper's limitations.

  A scholarly communication system is supposed to serve five functions, but can do so in a narrow sense as it currently does, or in a wider sense. These functions of the scholarly communication system are (1) registration-, (2) certification-, (3) awareness-, and (4) archival [@roosendaal1997], and (5) incentives [@10.1045/september2004-vandesompel]. A narrow fulfillment of for example the registration function would mean that findings that are published are registered, but not all findings are registered [e.g., due to selective publication; @10.1126/science.1255484]. Similarly, certification is supposed to occur through peer review, but peer review can exacerbate human biases in the assessment of quality [e.g., statistical significance increasing the perceived quality of methods; @10.1007/bf01173636].

  We propose an alternative design for scholarly communication based on piecemeal research outputs with direct links between subsequent outputs, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, we propose to  communicate piecemeal parts of the research cycle on an "as-you-go" basis. These pieces could be similar to sections of a research paper, but extend to research outputs such as software or materials. An "as-you-go" communication design respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

  With a network structure between piecemeal outputs of knowledge, we can go beyond citations and facilitate different questions about single- or collectives of knowledge. For example, how central is a single output in the larger network? Or: How densely interconnected is this collective of knowledge outputs? A network could facilitate question-driven evaluation where an indicator needs to be operationalized per question, instead of indicators that have become a goal in themselves and become invalidated by clear cheating behaviors [@Seeber2017; @PLOSMedicine2006]. As such, we propose to make evaluation of research its own research process with question formulation, operationalizations, and data collection (i.e., constructing the network of interest).

@@

  When stages of research are communicated separately and as they occur, it changes the communicative design to eliminate hindsight and allows more types of outputs to be communicated. For example, a theory can be communicated first and hypotheses communicated afterwards, as a direct descendant of the theory. Subsequently, a design can be linked as a direct descendant of the hypotheses, materials as a direct descendant of the design, and so on. This would allow for the incorporation of materials, data, and analysis code (amongst others). In this structure, many nodes could link to a single node (e.g., replication causes many data nodes to connect to the same hypotheses node) but one node can also link to many other nodes (e.g., when hypotheses follow from multiple theories or when a meta-analytic node is linked to many results nodes).

  Figure 1 shows a simple example of how these different research outputs (i.e., nodes) would directly connect to each other. The connection between these nodes only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion section could be a direct descendant of a results section and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last node, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

@@

  Given that these piecemeal outputs would be communicated as they occur, chronology is directly embedded in the communication process with many added benefits. For example, preregistration of hypotheses tries to ensure that predictions precede observations, which would be embedded with piecemeal communication where predictions are communicated when they are made [exploratory research could be communicated without hypotheses; for a more extensive discussion of the benefits and limits of preregistration see @10.1073/pnas.1708274114]. Moreover, if research outputs are communicated as they are produced, selective reporting (i.e., publication bias) is reduced by having already communicated the data before results are generated.

@@

  A continuous and network based communication system could take a wider interpretation of the scholarly functions it is supposed to serve [@roosendaal1997; @10.1045/september2004-vandesompel]. Registration would become more complete, because selective publication based on results is preempted by embedding communication before any results are known. Certification is improved by embedding the chronology of a research cycle into the communication of research, ensuring that predictions precede results [@10.1073/pnas.1708274114]. Awareness is improved by using open by design principles, whereas awareness is now limited by financial means to access scholarly articles [@10.12688/f1000research.12037.3]. Archival would not only be simplified with peer-to-peer protocols, but also allows anyone to create a copy and could result in excessive redundancy under the Lots Of Copies Keeps Stuff Safe principle [LOCKSS; @10.1045/june2001-reich]. In the next sections, we extend on how incentives could be adjusted in such a network structure, to facilitate both the evaluation of research(ers) and the planning of research.

@@

  The current scholarly communication system based on research articles is “after-the-fact” and can be supplemented by an “as-you-go” based communication system. By doing so, the functions of a scholarly communication system can be interpreted more widely, making registration complete, certification part of the process instead of just the judgment of peers, access to everything for everyone based on peer-to-peer protocols, simplify archival, and facilitate incentive structures that could align researcher’s interests with that of scholarly research.

>>>>

  Scholarly research faces threats to its sustainability and has been said to face a reproducibility crisis [@Baker2016] amongst other pernicious problems such as access and exclusivity. The underlying cause might be the way we have collectively designed the reporting and rewarding of research (implicitly or explicitly). The current scholarly communication system is primarily organized around researchers who publish static (digital) research papers in scholarly journals. Many of these journals have artificial page limits (in the digital age), which leads to artificial scarcity and subsequently increases the perceived prestige of such a journal due to high rejection rates (71% on average for APA journals in 2016; [perma.cc/Q7AT-RN5C](https://perma.cc/Q7AT-RN5C)). Furthermore, scholarly communication has become highly centralized, where over 50% of all papers are published by as little as five publishers (over 70% for social sciences) [@Lariviere2015]. Centralization has introduced knowledge discrimination, as publishers are able to influence who can access scholarly knowledge, what gets published, and allows for other single points of failure to arise with their own consequences (e.g., censorship; [perma.cc/HDX8-DJ8F](https://perma.cc/HDX8-DJ8F)). In order to have a sustainable scholarly research system, we consider it necessary to implement changes that provide progress on multiple of these threats at once instead of addressing them individually.

  Systems design directly affects what the system and the people who use it can do; scholarly communication still retains an analog based design affecting the effectivity of the spread and production of knowledge dissemination (see also [@Kling2005]). Researchers and institutions are evaluated on where and how much papers they publish (as a form of prestige). For example, an oft-used measure of quality is the Journal Impact Factor (JIF) [@Garfield2006]. The JIF is also frequently used to evaluate the 'quality' of individual papers under the assumption that a high impact factor predicts the success of individual papers (this assumption has been debunked many times) [@Prathap2016; @Seglen1992; @Seglen1994]. Many other performance indicators in the current system (e.g., citation counts and h-indices) resort to generic bean counting. Inadequate evaluation measures leave universities, individual researchers, and funders (amongst others) in the dark with respect to the substantive questions they might have about the produced scholarly knowledge. Additionally, work that is not aptly captured by the authorship of papers is likely to receive less recognition (e.g., writing software code) due to reward systems counting publications instead of contributions (see also [perma.cc/MUH7-VCA9](https://perma.cc/MUH7-VCA9)). It is unfeasible that a paper-based approach to scholarly communication can escape the consequences of paper's limitations.

  A scholarly communication system is supposed to serve five functions, but can do so in a narrow sense as it currently does, or in a wider sense. These functions of the scholarly communication system are (1) registration-, (2) certification-, (3) awareness-, and (4) archival [@roosendaal1997], and (5) incentives [@VandeSompel2004]. A narrow fulfillment of for example the registration function would mean that findings that are published are registered, but not all findings are registered (e.g., due to selective publication; [@10.1126/science.1255484]). Similarly, certification is supposed to occur through peer review, but peer review can exacerbate human biases in the assessment of quality (e.g., statistical significance increasing the perceived quality of methods; [@10.1007/bf01173636]).

  We propose an alternative design for scholarly communication based on modular research outputs with direct links between subsequent modules, forming a network. Whereas a paper-based approach communicates after a whole research cycle is completed, modular communication was proposed two decades ago [@Kircz1998;@VandeSompel2004;@10.7717/peerj-cs.78;@10.3233/ISU-2010-0613;@10.1163/095796511X560006]. These modules could be similar to sections of a research paper, but extend to modular research outputs such as software or materials. We propose to implement this modular communication on an "as-you-go" basis and include direct links to indicate provenance. This respects the chronological nature of research cycles and decreases the possibility for pernicious problems such as selective publication and making predictions after results are known (HARKing) [@Kerr1998].

  With a network structure between modules of knowledge, we can go beyond citations and facilitate different questions about single- or collectives of knowledge. For example, how central is a single module in the larger network? Or: How densely interconnected is this collective of knowledge modules? A network could facilitate question-driven evaluation where an indicator needs to be operationalized per question, instead of indicators that have become a goal in themselves and become invalidated by clear cheating behaviors [@Seeber2017; @PLOSMedicine2006]. As such, we propose to make evaluation of research its own research process with question formulation, operationalizations, and data collection (i.e., constructing the network of interest).

@@

  When stages of research are communicated separately and as they occur, it changes the communicative design to eliminate hindsight and allows more types of outputs to be communicated as separate modules. For example, a theory can be communicated first and hypotheses communicated afterwards, as a direct descendant of the theory. Subsequently, a design can be linked as a direct descendant of the hypotheses, materials as a direct descendant of the design, and so on. This would allow for the incorporation of materials, data, and analysis code (amongst others). In this structure, many nodes could link to a single node (e.g., replication causes many data nodes to connect to the same hypotheses node) but one node can also link to many other nodes (e.g., when hypotheses follow from multiple theories or when a meta-analytic node is linked to many results nodes).

  Figure 1 shows a simple example of how these different modular research outputs (i.e., nodes) would directly connect to each other. The connection between these nodes only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion section could be a direct descendant of a results section and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last node, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

@@

  Given that these modular outputs would be communicated as they occur, chronology is directly embedded in the communication process with many added benefits. For example, preregistration of hypotheses tries to ensure that predictions precede observations, which would be embedded with piecemeal communication where predictions are communicated when they are made [exploratory research could be communicated without hypotheses; for a more extensive discussion of the benefits and limits of preregistration see @10.1073/pnas.1708274114]. Moreover, if modular research outputs are communicated as they are produced, selective reporting (i.e., publication bias) is reduced by having already communicated the data before results are generated.

@@

  A continuous and network based communication system could take a wider interpretation of the scholarly functions it is supposed to serve [@roosendaal1997; @10.1045/september2004-vandesompel]. Registration would become more complete, because selective publication based on results is preempted by embedding communication before any results are known. Certification is improved by embedding the chronology of a research cycle into the communication of research, ensuring that predictions precede results [@10.1073/pnas.1708274114]. Awareness is improved by using open by design principles, whereas awareness is now limited by financial means to access scholarly papers [@10.12688/f1000research.12037.3]. Archival would not only be simplified with peer-to-peer protocols, but also allows anyone to create a copy and could result in excessive redundancy under the Lots Of Copies Keeps Stuff Safe principle [LOCKSS; @10.1045/june2001-reich]. In the next sections, we extend on how incentives could be adjusted in such a network structure, to facilitate both the evaluation of research(ers) and the planning of research.

@@

  The current scholarly communication system based on research papers is “after-the-fact” and can be supplemented by an “as-you-go” based communication system. By doing so, the functions of a scholarly communication system can be interpreted more widely, making registration complete, certification part of the process instead of just the judgment of peers, access to everything for everyone based on peer-to-peer protocols, simplify archival, and facilitate incentive structures that could align researcher’s interests with that of scholarly research.
